\myfont Implementation and evaluation of register allocation for translating LLVM-{}- to x86-64

### 2024-01-14 article.tex
Fix code offset

\begin{figure}[H]
     \centering
     \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
let rec linearscan insns spills ?(v = 0)
    (avail, active, assign, incstart, incend) =
  match IT.min_binding_opt incstart with
  | Some (i, ss) ->
      let e = S.SS.choose ss in
      let ss' = S.SS.remove e ss in
      let incstart' =
        if S.SS.is_empty ss' then IT.remove i incstart
        else IT.add i ss' incstart
      in
      let spills, (avail', active', assign', incstart', incend') =
        match Regs.choose_opt avail with
        | Some reg ->
            ( spills,
              expire i
                ( Regs.remove reg avail,
                  S.ST.add e reg active,
                  S.ST.add e (Reg reg) assign,
                  incstart',
                  incend ) )
        | None ->
            let _j, spill =
              IT.filter
                (fun _k v -> S.SS.exists (fun e -> S.ST.mem e active) v)
                incend
              |> IT.max_binding
            in
            let spill =
              S.SS.filter (fun e -> S.ST.mem e active) spill |> S.SS.choose
            in
            let reg = S.ST.find spill active in
            let incend' =
              IT.filter_map
                (fun _ ss ->
                  let ss' = S.SS.remove spill ss in
                  if S.SS.is_empty ss' then None else Some ss')
                incend
            in
     \end{minted}
\end{figure}

\begin{figure}[H]
     \centering
     \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
            ( spills + 1,
              expire i
                ( avail,
                  S.ST.remove spill active |> S.ST.add e reg,
                  S.ST.remove spill assign
                  |> S.ST.add spill (stack spills)
                  |> S.ST.add e (Reg reg),
                  incstart',
                  incend' ) )
      in
      linearscan insns spills (avail', active', assign', incstart', incend')
  | None -> expire (List.length insns) (avail, active, assign, incstart, incend)
     \end{minted}
     \caption{Implementation of \texttt{linearscan} in \texttt{linear.ml}}\label{fig:linearscan-ocaml}
\end{figure}
\noindent The implementation of the actual linearscan iteration in Figure \ref{fig:linearscan-ocaml} is quite tedious, but the core idea is the outerermost \texttt{match} block. The smallest interval is selected using \texttt{min\_binding\_opt}. This is done to visit each interval in increasing order of start point. If there are none left, the linearscan iteration is essentially over, but calls \texttt{expire} regardless to clean things up. The case of intervals remaining is more interesting, as one of the variables starting at \(i\) needs to be assigned. First a variable is chosen, the order of this is not important. Despite the minimal example used in Figure \ref{fig:linear.ll}, there are cases where several intervals begin at the same point (e.g. loops), hence a symbol set is used. Lines 7-10 are just concerned with removing the chosen variable from the table of increasing start points. Then, an available register is chosen. If one is available (line 13), it is 1) removed from the set of available registers, 2) added to the symbol table of active variables, and 3) assigned the same register in the returned assignments table. If no registers are available on the other hand (line 21), a variable needs to be spilled. It does this by first filtering through and only visiting the end points that are currently active (i.e. members of \texttt{active}). Then the one of intervals that are currently active with the greatest end points is chosen by first filtering end points for which at least of the variables is active (lines 23-26), and then selecting one of those immediately after (line 29). The currently assigned register of that variable is then looked up (line 31) and the end point is removed (lines 33-37).

>>>>

\begin{figure}[H]
     \centering
     \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
let rec linearscan insns spills (avail, active, assign, incstart, incend) =
  match IT.min_binding_opt incstart with
  | Some (i, ss) ->
      let e = S.SS.choose ss in
      let ss' = S.SS.remove e ss in
      let incstart' =
        if S.SS.is_empty ss' then IT.remove i incstart
        else IT.add i ss' incstart
      in
      let spills, (avail', active', assign', incstart', incend') =
        match Regs.choose_opt avail with
        | Some reg ->
            ( spills,
              expire i
                ( Regs.remove reg avail,
                  S.ST.add e reg active,
                  S.ST.add e (Reg reg) assign,
     \end{minted}
\end{figure}

\begin{figure}[H]
     \centering
     \begin{minted}[linenos,bgcolor=LightGray,firstnumber=18]{ocaml}
                  incstart',
                  incend ) )
        | None ->
            let _j, spill =
              IT.filter
                (fun _k v -> S.SS.exists (fun e -> S.ST.mem e active) v)
                incend
              |> IT.max_binding
            in
            let spill =
              S.SS.filter (fun e -> S.ST.mem e active) spill |> S.SS.choose
            in
            let reg = S.ST.find spill active in
            let incend' =
              IT.filter_map
                (fun _ ss ->
                  let ss' = S.SS.remove spill ss in
                  if S.SS.is_empty ss' then None else Some ss')
                incend
            in
            ( spills + 1,
              expire i
                ( avail,
                  S.ST.remove spill active |> S.ST.add e reg,
                  S.ST.remove spill assign
                  |> S.ST.add spill (stack spills)
                  |> S.ST.add e (Reg reg),
                  incstart',
                  incend' ) )
      in
      linearscan insns spills (avail', active', assign', incstart', incend')
  | None -> expire (List.length insns) (avail, active, assign, incstart, incend)
     \end{minted}
     \caption{Implementation of \texttt{linearscan} in \texttt{linear.ml}}\label{fig:linearscan-ocaml}
\end{figure}
\noindent The implementation of the actual linearscan iteration in Figure \ref{fig:linearscan-ocaml} is quite tedious, but the core idea is the outerermost \texttt{match} block. The smallest interval is selected using \texttt{min\_binding\_opt}. This is done to visit each interval in increasing order of start point. If there are none left, the linearscan iteration is essentially over, but calls \texttt{expire} regardless to clean things up. The case of intervals remaining is more interesting, as one of the variables starting at \(i\) needs to be assigned. First a variable is chosen, the order of this is not important. Despite the minimal example used in Figure \ref{fig:linear.ll}, there are cases where several intervals begin at the same point (e.g. loops), hence a symbol set is used. Lines 7-10 are just concerned with removing the chosen variable from the table of increasing start points. Then, an available register is chosen. If one is available (line 13), it is 1) removed from the set of available registers, 2) added to the symbol table of active variables, and 3) assigned the same register in the returned assignments table. If no registers are available on the other hand (line 21), a variable needs to be spilled. It does this by first filtering through and only visiting the end points that are currently active (i.e. members of \texttt{active}). Then the one of intervals that are currently active with the greatest end points is chosen by first filtering end points for which at least of the variables is active (lines 23-26), and then selecting one of those immediately after (line 29). The currently assigned register of that variable is then looked up (line 31) and the end point is removed (lines 33-37).
